---
title: "Exercise 11"
author: "Maya Eldani"
format: html
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(skimr)
library(naniar)
library(MuMIn)
```

## Load the data

```{r}
d <- read_tsv("https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt")
```

---

## Data Cleaning

```{r}
# Replace -999 with NA
d <- d %>%
  mutate(across(where(is.numeric), ~na_if(., -999)))

# Drop unnecessary columns and rename
d <- d %>%
  select(-`litter size`, -refs) %>%
  rename(
    `maxlife(mo)` = `max. life(mo)`,
    `weanmass(g)` = `wean mass(g)`
  )

# Log transform
d <- d %>%
  mutate(across(where(is.numeric), log))
```

---

## Rename variables for modeling

```{r}
d <- d %>%
  rename(
    body_mass = `mass(g)`,
    gestation = `gestation(mo)`,
    weaning = `weaning(mo)`,
    afr = `AFR(mo)`,
    maxlife = `maxlife(mo)`,
    newborn = `newborn(g)`,
    weanmass = `weanmass(g)`
  )
```

---

## Step 4: Regression and Residuals

```{r}
m_gest <- lm(gestation ~ body_mass, data = d, na.action = na.exclude)
m_wean <- lm(weaning ~ body_mass, data = d, na.action = na.exclude)
m_afr  <- lm(afr ~ body_mass, data = d, na.action = na.exclude)
m_life <- lm(maxlife ~ body_mass, data = d, na.action = na.exclude)

m_newborn  <- lm(newborn ~ body_mass, data = d, na.action = na.exclude)
m_weanmass <- lm(weanmass ~ body_mass, data = d, na.action = na.exclude)

d <- d %>%
  mutate(
    relGest = resid(m_gest),
    relWean = resid(m_wean),
    relAFR = resid(m_afr),
    relLife = resid(m_life),
    relNewbornMass = resid(m_newborn),
    relWeaningMass = resid(m_weanmass)
  )
```

### Interpretation

These residuals represent life history traits relative to body mass. Positive values indicate higher-than-expected values for a given body size, while negative values indicate lower-than-expected values.

---

## Step 5: Residual Plots

```{r}
ggplot(d, aes(x = reorder(order, relLife, median, na.rm = TRUE), y = relLife)) +
  geom_boxplot() +
  coord_flip() +
  theme_minimal()
```

```{r}
ggplot(d, aes(x = reorder(order, relNewbornMass, median, na.rm = TRUE), y = relNewbornMass)) +
  geom_boxplot() +
  coord_flip() +
  theme_minimal()
```

```{r}
ggplot(d, aes(x = reorder(order, relWeaningMass, median, na.rm = TRUE), y = relWeaningMass)) +
  geom_boxplot() +
  coord_flip() +
  theme_minimal()
```

---

## Step 6: Model Selection (maxlife)

```{r}
d_life <- d %>%
  select(maxlife, gestation, newborn, weaning, weanmass, `litters/year`, body_mass) %>%
  drop_na()

options(na.action = "na.fail")

m_global_life <- lm(maxlife ~ gestation + newborn + weaning + weanmass +
                      `litters/year` + body_mass,
                    data = d_life)

ms_life <- dredge(m_global_life)

top_life <- subset(ms_life, delta <= 4)
nrow(top_life)

models_life <- get.models(ms_life, subset = delta <= 4)

avg_life <- model.avg(models_life)

summary(avg_life)
confint(avg_life)
```

### Interpretation

The best model is the one with the lowest AICc value. Models with ΔAICc ≤ 4 are considered competitive. Variables that appear across these models are the most consistent predictors of lifespan.

---

## Step 7: Model Selection (AFR)

```{r}
d_afr <- d %>%
  select(afr, gestation, newborn, weaning, weanmass, `litters/year`, body_mass) %>%
  drop_na()

m_global_afr <- lm(afr ~ gestation + newborn + weaning + weanmass +
                     `litters/year` + body_mass,
                   data = d_afr)

ms_afr <- dredge(m_global_afr)

top_afr <- subset(ms_afr, delta <= 4)
nrow(top_afr)

models_afr <- get.models(ms_afr, subset = delta <= 4)

avg_afr <- model.avg(models_afr)

summary(avg_afr)
confint(avg_afr)
```
